Predicting Tags for StackOverflow Posts

### 2013-06-19 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: badac33fe9e9bccc490f1e6696b25811990b7c37 [formerly 35a9c0ad732c4977e428241535871cd85af3cd34]
Former-commit-id: 625efd9cee1c3704adb831523ccfb508bfb012ac
ICCM revisions
Former-commit-id: e5687596b15d3449f5506054428777fb51b0c944 [formerly f7232924e39102cbc7c1081c9846ad260e866bbd]
Former-commit-id: e8cc0b891f6b2bd7707de7aaf7acb88c2ccb70b2

\begin{abstract}
  Hashtags created by authors of online content provide a view of a user's goals and interests.
  Predicting users' interests can lead to improved, more user-centered human-computer systems.
  Large-scale behavioral datasets such as Twitter, Facebook, LinkedIn, StackOverflow, and Yelp can be mined and explored to study these hashtags and how they relate to users' interests.
  We explored the StackOverflow dataset, and developed an ACT-R inspired Bayesian probabilistic model that can predict the hashtags used by the author of the post.
  The model is 65\% accurate when tasked to predict one tag per post on average.
  This is achieved by choosing the tag that has the highest log odds of being correct,
  given the tag's prior log odds of occurrence and adjusting for the log likelihood ratio of the words in the post being associated with the tag.
  The model is a successful case showing that ACT-R's declarative memory retrieval equations scale, and are relevant to task domains that require large-scale knowledge stores.

@@

Predicting tags for the StackOverflow dataset has been worked on by at least one prior researcher \cite{Kuo2011}.
\citeauthor{Kuo2011} uses a co-occurrence model that predicts tags based on the words in the post and their relation (co-occurrence) to tags.
The model was initially built for next-word prediction in large documents, and then adapted to the StackOverflow dataset by constraining the next word predicted to only tags.
His co-occurrence model has a 47\% classification accuracy for predicting StackOverflow tags when the model predicts one tag per post.

The SNIF-ACT \cite{Fu2007} model also uses co-occurrences and can predict link traversal for search queries.
That is, given a search query (goal state), fetched results, and words included in those fetched results, the model predicts the most likely link that a person will click on.
This StackOverflow model is similar because it leverages ACT-R's declarative memory retrieval mechanisms to generate predictions (tag activations instead of link activations).
However, it is slightly different because it takes into account prior odds of tag activations which was not necessary in SNIF-ACT.

@@

A variation of ACT-R's declarative memory retrieval theory \cite{Anderson2004} was used to model tag retrievals for StackOverflow posts.
The strength of association between post words and tags was calculated using 2/3 of the dataset (one million posts).
A logistic regression statistical technique was used to calibrate model parameters and optimize performance.
The weights for each model component were calibrated using a sample of \num{1000} posts not contained within the strength of association dataset.
The calibrated model was then tested on a new sample of \num{1000} posts not contained within either previous training dataset.
Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).

@@

For a given post, the model produces a vector of tag activations.
For that post, the author's chosen tags are known (see Figure \ref{fig:examplePost}).
So model performance can be measured by comparing the model's tag activations with chosen tags for a post, and then aggregated across posts to get an overall average performance.

@@

\subsection{Logistic Regression}

@@

The final best-fit coefficients for each model component and formal statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits} respectively.
Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.52$, npv = .994, ppv = .70).
It is also interesting that although the attentional weighting ($W$) terms for the title and body were calibrated by the regression, they did not stray far from cognitively-plausible defaults
(see coefficients for title and body in Table \ref{tab:coeffs}).

\subsection{Example Posts}

A visualization of the model's tag activation for the example post in Figure \ref{fig:examplePost} is included in Figure \ref{fig:modelPost}.

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-act-crop.pdf}}
  \caption{
    Model's prediction of top-ten tags for post in Figure \ref{fig:examplePost}.
    The title is the title text of the post.
    The attentional weight (scaled entropy) for each word in the title is included below each word.
    The user-chosen tags are included below the attentional weights.
    The stacked bar graphs are the total activation for the top-ten predicted model tags for the post.
    Each bar graph is partitioned into the three model components that vary across the tags in a single post (not the offset).
}
  \label{fig:modelPost}
\end{figure}

Note how the base level activation and contextual activation combine to predict tags that are both frequently used and contextually relevant.
For example, note that the \emph{javascript} tag leverages a combination of contextual activation and prior activation to achieve a high total activation.
Alternatively, the \emph{svg} tag relies almost entirely on context to achieve high activation.
And note that the model is by no means perfect.
For example, \emph{html} and \emph{jquery} are not tagged by the author, but the model ranks them as the second and third highest tag.
And the author's tag \emph{firefox} is only ranked 12th by the model.

\subsection{Generalizability}

@@

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{ROC-crop.pdf}}
  \caption{
    Model performance across the calibration and test dataset, as well as when specific model components are removed.
    The y-axis represents model performance (one is perfect), and the x-axis is the proportion of tags that the model generates compared to the number of tags used by the authors of the posts (one is equal number).
    The vertical dotted line represents the point where the model is asked to tag on average one tag per post.
  }
  \label{fig:ROC}
\end{figure}

@@

\subsection{Necessity Analysis}

Figure \ref{fig:ROC} also shows model performance after specific model components are removed.
Model parameters were recalibrated for each model, except for the full model test dataset that used the parameters from the calibration dataset.
These plots provide useful information regarding the relative necessity of each component in the full model.
That is, which model components contribute most strongly to increased performance, and which model components might be redundant (not necessary).
Note that it is certainly the case that each model component contributes positive and unique variance to model performance.
However, removing only one component (particularly the words in the body of the post) does not drastically reduce model performance.
It is only when groups of components are removed that model performance starts to drastically suffer.

@@

The model could also be used as a tag cleanup system for stackoverflow.com.
In this scenario the tags for a previously submitted post would be compared against model predictions.
If the model's predicted tag activation for a used tag is substantially low, then that tag could either be removed from the post, or the post could be flagged for further analysis.

@@

Our model achieves 65\% classification accuracy when predicting one tag per post on average
(height of full model at vertical dotted line in Figure \ref{fig:ROC})
compared to 47\% accuracy for the co-occurrence model presented in \citeA{Kuo2011}.
Most likely we achieved the increase in model performance by adding some additional model components:
The offset and possibly the body term as well,
as it was unclear if \citeauthor{Kuo2011} used context from both the body and title of a post or just the title.
Also we used two orders of magnitude more posts to train the model co-occurrence matrix ($S_{ji}$).
Nonetheless, there are certainly model similarities:
Co-occurrence count was used as the core measure of similarity, and both local (strength of association) and global (entropy or idf) weighting measures were used to compute tag likelihood, given context.

@@

@@

Alongside exploring the offset component of the model, our main planned research is to include the post author's past tagging behavior as an additional model predictor.
We are exploring a Bayesian updating technique that would adjust prior tag activations from global priors to user's priors as more tagging information about the user is collected.
For the StackOverflow dataset, adding user's tagging history to the model should improve predictive power for veteran users,
since programmers tend to specialize on a few set of tags that are of interest to them.
This technique is possible, since a post author must be signed in before creating a post.

@@

We would like to thank Fred Oswald for his excellent guidance and recommendations throughout the model development and data analysis process.

>>>>

\begin{abstract}
  Hashtags created by authors of online content provide a view of a user's goals and interests.
  Predicting users' interests can lead to improved, more user centered human-computer systems.
  Large-scale behavioral datasets such as Twitter, Facebook, LinkedIn, StackOverflow, and Yelp can be mined and explored to study these hashtags and how they relate to users' interests.
  We explored the StackOverflow dataset, and developed an ACT-R inspired Bayesian probabilistic model that can predict the hashtags used by the author of the post.
  The model is 65\% accurate when tasked to predict one tag per post on average.
  This is achieved by choosing the tag that has the highest log odds of being correct,
  given the tag's prior log odds of occurrence and adjusting for the log likelihood ratio of the words in the post being associated with the tag.
  The model is a successful case showing that ACT-R's declarative memory retrieval equations scale, and are relevant to task domains that require large-scale knowledge stores.

@@

At least one prior researcher \cite{Kuo2011} has worked on StackOverflow tag prediction.
\citeauthor{Kuo2011} uses a co-occurrence model that predicts tags based on the words in the post and their relation (co-occurrence) to tags.
The model was initially built for next-word prediction in large documents, and then adapted to the StackOverflow dataset by constraining the next word predicted to only tags.
His co-occurrence model has a 47\% classification accuracy predicting one tag per post.

The SNIF-ACT \cite{Fu2007} model also uses co-occurrences and can predict link traversal for search queries.
That is, given a search query (goal state) and fetched results, the model predicts the most likely link that a person will click on.
This StackOverflow model is similar because it leverages ACT-R's declarative memory retrieval mechanisms to generate predictions (tag activations instead of link activations).
However, it is slightly different because it takes into account prior odds of tag activations which was not necessary in SNIF-ACT.

@@

A variation of ACT-R's declarative memory retrieval theory \cite{Anderson2004} was used to model tag retrievals for StackOverflow posts.
The strength of association between post words and tags was calculated using 2/3 of the dataset (one million posts).
A logistic regression statistical technique was used to calibrate model parameters and optimize performance.
The weights for each model component were calibrated using a sample of \num{1000} posts not contained within the strength of association dataset.
The calibrated model was then tested on a fresh sample of \num{1000} posts.
Classification accuracy was used as a metric of model performance (i.e., number of correctly-predicted tags).

@@

For a given post, the model produces a vector of tag activations.
For that post, the author's chosen tags are known (see Figure \ref{fig:examplePost}).
So model performance can be measured by comparing the model's tag activations with chosen tags for a post, and then aggregated across posts to get an average.

@@

\subsubsection{Logistic Regression}

@@

The final best-fit coefficients for each model component and statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits}.
All four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.52$, npv = .994, ppv = .70).
It is also interesting that although the attentional weighting ($W$) terms for the title and body were calibrated by the regression, they did not stray far from cognitively-plausible defaults
(see coefficients for title and body in Table \ref{tab:coeffs}).

\subsubsection{Example Posts}

A visualization of the model's tag activation for the post in Figure \ref{fig:examplePost} is included in Figure \ref{fig:modelPost}.

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-act-crop.pdf}}
  \caption{
    Model's prediction of the top tags for post in Figure \ref{fig:examplePost}.
    At the top is the title text of the post, then the attentional weight (scaled entropy) for each word in the title, then the user-chosen tags.
    The bar graphs are the total activation for the top predicted model tags for the post.
    Each bar graph is partitioned into the three model components that vary across the tags in a single post.
}
  \label{fig:modelPost}
\end{figure}

The base level activation and contextual activation combine to predict tags that are both frequently used and contextually relevant.
For example, note that the \emph{javascript} tag leverages a combination of contextual activation and prior activation to achieve a high total activation.
Alternatively, the \emph{svg} tag relies almost entirely on context to achieve high activation.
And note that the model is by no means perfect.
For example, \emph{html} and \emph{jquery} are not tagged by the author, but the model ranks them as the second and third highest tag.
And the author's tag \emph{firefox} is only ranked 12th by the model.

\subsubsection{Generalizability}

@@

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{ROC-crop.pdf}}
  \caption{
    Model performance across the calibration and test dataset, as well as when specific model components are removed.
    The y-axis is model performance (one is perfect), and the x-axis is the proportion of tags that the model generates compared to the total number of tags used by the post authors (one is equal number).
    The vertical dotted line is where the model generates one tag per post on average.
  }
  \label{fig:ROC}
\end{figure}

@@

\subsubsection{Necessity Analysis}

Figure \ref{fig:ROC} also shows performance after specific model components are removed.
Model parameters were recalibrated for each model, except for the full model test dataset that used the parameters from the calibration dataset.
These plots provide useful information regarding the relative necessity of each component in the full model.
That is, which model components contribute most strongly to increased performance, and which model components might be redundant (not necessary).
Note that it is certainly the case that each model component contributes positive and unique variance to model performance.
However, removing only one component (particularly the words in the body of the post) does not drastically reduce model performance.
It is only when groups of components are removed that model performance starts to drastically suffer.

@@

The model could also be used as a tag cleanup system.
In this scenario the tags for a previously submitted post would be compared against model predictions.
If the model's predicted tag activation for a used tag is substantially low, then that tag could either be removed from the post, or the post could be flagged for further analysis.

@@

Our model achieves 65\% classification accuracy when predicting one tag per post on average (see Figure \ref{fig:ROC})
compared to 47\% accuracy for the co-occurrence model presented in \citeA{Kuo2011}.
Most likely we achieved the increase in model performance by adding the offset model component and possibly the body term as well, 
as it was unclear if \citeauthor{Kuo2011} used context from both the body and title of a post or just the title.
Also we used two orders of magnitude more posts to train the model co-occurrence matrix ($S_{ji}$).
Nonetheless, there are certainly model similarities:
Co-occurrence count was used as the core measure of similarity, and both local (strength of association) and global (entropy or idf) weighting measures were used to compute tag likelihood, given context.

@@

\subsubsection{Model Plausibility}

ACT-R's declarative memory retrieval process is the core of the StackOverflow model.
This is a more constrained and cognitively-plausible model than the set of machine learning techniques used by \citeA{Kuo2011} for the same task. 
Interestingly, when using ACT-R's retrieval mechanism, the model performs significantly better than the less constrained models in \citeA{Kuo2011}.
So there is no reason to think that people do not engage in a process much like this to select tags.
However, more research is necessary before making the strong claim that the StackOverflow model \emph{is} the process people go through when selecting tags.
In particular, the tag errors made by authors should be looked at, and model behavior should be tested against these important counter examples.
Also, places where the StackOverflow model deviate from ACT-R's default retrieval theory (i.e., the offset term, entropy and attentional weighting) should be looked at in greater detail.

@@

Alongside exploring the offset component and testing cognitive plausibility, our main planned research is to include the post author's past tagging behavior as an additional model predictor.
We are exploring a Bayesian updating technique that would adjust prior tag activations from global priors to user's priors as more tagging information about the user is collected.
For the StackOverflow dataset, adding user's tagging history to the model should improve predictive power for veteran users,
since programmers tend to specialize on a few set of tags that are of interest to them.

@@

We would like to thank Fred Oswald for his excellent guidance and feedback throughout the model development process.
